A Continual Planner that Switches Between Serial and Decision-Theoretic Planning

### 2010-11-17 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Forgot the bibliography last time.
 - Charles

We are concerned with problems that feature partial
observability. These have a perceptual model given in terms of a
finite set of stochastic {\em senses} $\stochSenses$, deterministic
sensing outcomes $\detSenses$, and perceptual propositions
$\percepts$, called {\em percepts}. Here, an observation $\observ$ is
a set of percepts $\observ \subseteq \percepts$, and we denote
\observations\ the set of reachable observations. The underlying state
of the process cannot be observed directly, rather, senses
$\stochSense \in \stochSenses$ effect an observation $\observ \in
\observations$ that informs what should be believed about the state a
process is in. In detail, if $\action$ is applied effecting a
transition to a successor state $\state'$, then an observation occurs
according to the active senses $\stochSenses(\stochAction, \state')
\subseteq \stochSenses$. A precept is active, written $\stochSense \in
\stochSenses(\stochAction, \state')$, if the senses'
action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
\subseteq \propositions$ is satisfied by the state $\state'$; In the
usual sense that $\poss_\states(\stochSense) \subseteq \state'$.
%%
When a sense is active, nature must choose exactly one outcome amongst
a small set of deterministic choices $\detSenses(\stochSense)
\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
$i$ we have $\detSense_i \subseteq \percepts$. The probability of
the $i^{th}$ element being chosen is given by
$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
1$. The observation received by the agent corresponds to the union of
perceptual propositions from chosen elements of active senses.

A POMDP has a starting configuration that corresponds to a Bayesian
belief-state. Intuitively, this is the robot's subjective belief about
its environment. Technically, a belief-state $\bstate$ is a
probability distribution over process states. We write
$\bstate(\state)$ to denote the probability that the process is in
$\state$ according to $\bstate$, and $\bstate_0$ when discussing the
starting configuration. We suppose the latter is given in a factored
tree format.

@@

The modelling language of choice for planning in probabilistic
problems is the Probabilistic Planning Domain Definition
Language~\cite{younes:littman:04}. PPDDL was used in all 3 of the
International Planning Competitions since 2004. A variation on PDDL
for describing domains with stochastic actions and uncertain starting
configurations, PPDDL is a declarative first-order (a.k.a.,
relational) that facilitates factored description of domains and
problem instances. There are straightforward compilations from
problems expressed compactly in PPDDL to propositional representations
amenable to state-of-the-art planning procedures.  Because this
language cannot model domains that feature partial observability, for
our setting we have developed DTPDDL, an extension of PPDDL to partial
observability.

@@

\bibliography{papers}
\bibliographystyle{aaai}
\end{document}

>>>>

We are concerned with problems that feature partial
observability. These have a perceptual model given in terms of a
finite set of stochastic {\em senses} $\stochSenses$, deterministic
sensing outcomes $\detSenses$, and perceptual propositions
$\percepts$, called {\em percepts}. Here, an observation $\observ$ is
a set of percepts $\observ \subseteq \percepts$, and we denote
\observations\ the set of reachable observations. The underlying state
of the process cannot be observed directly, rather, senses
$\stochSense \in \stochSenses$ effect an observation $\observ \in
\observations$ that informs what should be believed about the state a
process is in. In detail, if $\action$ is applied effecting a
transition to a successor state $\state'$, then an observation occurs
according to the active senses $\stochSenses(\stochAction, \state')
\subseteq \stochSenses$. A percept is active, written $\stochSense \in
\stochSenses(\stochAction, \state')$, if the senses'
action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
\subseteq \propositions$ is satisfied by the state $\state'$; In the
usual sense that $\poss_\states(\stochSense) \subseteq \state'$.
%%
When a sense is active, nature must choose exactly one outcome amongst
a small set of deterministic choices $\detSenses(\stochSense)
\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
$i$ we have $\detSense_i \subseteq \percepts$. The probability of
the $i^{th}$ element being chosen is given by
$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
1$. The observation received by the agent corresponds to the union of
perceptual propositions from chosen elements of active senses.

A POMDP has a starting configuration that corresponds to a Bayesian
belief-state. Intuitively, this is the robot's subjective belief about
its environment. Formally, a belief-state $\bstate$ is a probability
distribution over process states. We write $\bstate(\state)$ to denote
the probability that the process is in $\state$ according to
$\bstate$, and $\bstate_0$ when discussing the starting
configuration.

@@

The modelling language of choice for planning in probabilistic
problems is the Probabilistic Planning Domain Definition
Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL was used in
all 3 of the International Planning Competitions since 2004. A
variation on PDDL for describing domains with stochastic actions and
uncertain starting configurations, PPDDL is a declarative first-order
(a.k.a., relational) that facilitates factored description of domains
and problem instances. There are straightforward compilations from
problems expressed compactly in PPDDL to propositional representations
amenable to state-of-the-art planning procedures.  Because this
language cannot model domains that feature partial observability, for
our setting we have developed DTPDDL, an extension of PPDDL to partial
observability.

We suppose $\bstate_0$ is given in a factored tree format of the
form:

\def\oom{$^{\tt +}$}
\def\zom{$^*$}
\def\bump{\hspace{1cm}}
\def\req#1{$^{\tt #1}$}
\def\noteme#1{}%{[[#1]]}
\def\notecoauth#1{\ }
\def\notereader#1{[[#1]]}
\def\meta{$\uparrow\uparrow$}
\def\la{\langle}
\def\ra{\rangle}

\scriptsize
\begin{nopagebreak}\begin{tabtt}
\bump \= <observation-def> \bump\=::= \=(:obs\=erve <o-symbol> \+\+\+\+\\
                        %%{[}<extension-def>]  \\
                        :parameters (<typed list (variable)>)  \\
                        <o-def body>) \-\-\- \\
  <o-def body> \> ::= [:state <emptyOr (pre-GD)>] \\
  \> \> [:action <atomic action(term)> ] \\
    \> \> [:effect <emptyOr (o-effect)>] \\
  <o-symbol> \> ::= <name> \\
  <atomic action(t)> \> ::= (<action-symbol> t\zom) \\
  <o-effect> \> ::= (and <c-o-effect>\zom) \\
  <o-effect> \> ::= <c-o-effect> \\
  <c-o-effect> \> ::= \req{:conditional-effects} (when <GD>
                        <o-effect>) \\  
  <c-o-effect> \> ::= \req{:universal-effects} (forall (<typed list (var)>) <o-effect>) \\
  <c-o-effect> \> ::= \req{:probabilistic-effects} (probabilistic
  <prob> <o-effect>) \\
  <c-o-effect> \> ::=  \req{:probabilistic-effects}\req{:universal-unwinding} \\
%% 
\hspace{3ex}(probabilistic (for-each (<typed list (x)>) <s-f-head>  <o-effect> )) \\
  <c-o-effect> \> ::= <p-o-effect> \\
  <atomic o-formula(t)> \> ::= (<observation> t\zom) \\
  <p-o-effect> \> ::= <atomic o-formula(term)> \\
  <p-o-effect> \> ::= (not <atomic o-formula(term)>) \\
  <p-o-effect> \> ::= \req{:fluents}(<assign-op> <o-f-head> <o-f-exp>) \\
  <o-f-comp> \> ::= (<binary-comp> <o-f-exp> <o-f-exp>)\\
  <o-f-exp> \> ::= <number>\\
  <o-f-exp> \> ::= (<binary-op>  <o-f-exp> <o-f-exp>)\\
  <o-f-exp> \> ::= (<multi-op>  <o-f-exp> <o-f-exp>\oom)\\
  <o-f-exp> \> ::= (- <o-f-exp>)\\
  <o-f-exp> \> ::= <o-f-head>\\
  <o-f-head> \> ::= (<o-function-symbol> <term>\zom )\\
  <o-f-head> \> ::= <o-function-symbol>\\
\end{tabtt}\end{nopagebreak}
\noteme{<structure-def> ::= <attach-def>}
\normalsize

%% \small{
%% \begin{tabtt}
%% (:init  (and \=f_{11} f_{12} ..
%%  \> (probabilistic p_{}) 
%%  \> (probabilistic ..) ..) \\
%% \end{tabtt}
%% }

%% \[(:\pp{init} (and f_{11} f_{12} .. (\pp{probabilistic} p_) ))\]

\noindent Nodes of the tree have one of three types: (1) atomic nodes
are labelled with a state proposition, e.g., $(=
(\pp{location}~\pp{cornflakes}) \pp{kitchen})$, (2) probabilistic
nodes, have the form $(r_1, C_1;..;r_n, C_n)$, and (3) constructive
nodes, take the form of a set of probabilistic and atomic nodes. The
root node is conjunctive, and leaves are always atomic. A convenient
way to give semantics to the tree, is to treat nodes as if they were
actions. An atomic node is an action with a positive effect
corresponding to the node label, .

@@

\bibliographystyle{aaai}
\bibliography{papers}

\end{document}
